Collaborative Test-Driven Scientific Model Validation

### 2013-11-23 sciunit-icse14/sciunit-icse14.tex
more progress on intro

Code review can be quite effective \cite{creffectiveness1,creffectiveness2}, but this requires that developers  expend considerable effort \cite{effort}. Code review is also most effective for resolving  issues related to software architecture \cite{evolvability}. 
Most development teams thus supplement code reviews with {automated} approaches to verification and validation, the most widely-used  of which is \emph{unit testing} \cite{beck2003}. In brief, unit tests are functions that check a single portion of a program against a single well-defined correctness criterion. A suite of such tests can be seen as a partial specification of a program or component.  Test suites complement code review by allowing developers to answer questions like these more easily:
\begin{enumerate}
\item Which functionality has been adequately implemented? What remains to be done?
%\item How much progress does an individual code contribution make? 
\item What modes of usage does the team consider most important? How does the team measure correctness?
\item Does a code contribution cause \emph{regressions} in other parts of a program?
\end{enumerate}
Scientists ask analagous questions:
\begin{enumerate}
\item Which observations are already well-explained by existing models? What is the state-of-the-art? What are the open modeling problems of interest?
%\item How much better is a new model than the state of the art? What are the open problems in my field?
\item What are the contemporary community standards for measuring goodness-of-fit?
\item How do newly-made experimental observations impact the validity of previous models?
\end{enumerate}
But while software engineers can rely on a program's test suite, scientists today must extract this information from a body of scientific publications. This is increasingly difficult. Each publication focuses on just one model and is frozen in time, so it does not consider the latest experimental data or statistical methods. Discovering, precisely characterizing and comparing several models to discover the state of the art and find open modeling problems often requires an encyclopedic knowledge of the literature. Senior scientists can attempt to summarize this by writing review papers, but in many areas, the number of publications generated every year can be overwhelming \cite{jinha_article_2010}, and comprehensive reviews are published relatively infrequently for any given area. Statisticians commonly complain that scientists are not following best practices and that community standards evolve too slowly because canonical papers or popular reviews advocate outdated methods \cite{grumpystatisticians}. One might compare this process to a software development team relying primarily on carefully-prepared summaries of code review sessions, peppered with cross-references to previous summaries, to understand and validate a complex software project.

This suggests that quantitative model validation could be improved by the adoption of test-driven methodologies alongside peer review. However, the scientific community presents several unique challenges:
\begin{enumerate}
\item Unit tests are typically pass/fail, while goodness-of-fit is typically a continuous value (e.g. a $p$-value). 
\item Different communities prefer different goodness-of-fit measures. These continuously evolve as, for example, statisticians develop better methods.
\item Model developers compete with one another to predict data, rather than collaborating with one another on a common codebase. They are sometimes reluctant to share certain details of their implementations \cite{badscientists}.
\item Experimental data can itself be erroneous, noisy or inconsistent with existing data.
\item Data formats differ, sometimes drastically, between experimentalists. Similarly, different modelers may prefer different programming languages or libraries.
\item Professional software developers are typically trained in testing practices and tools, while scientists  rarely use testing practices with their own code \cite{badscientists2}.
\end{enumerate}
%Anecdotally, one of the greatest fears of new scientists is that a problem that they are working on has already been solved and published.

>>>>

Code review can be quite effective \cite{creffectiveness1,creffectiveness2}, but this requires that developers  expend considerable effort \cite{effort}. Code review is also most effective for resolving  issues related to software architecture \cite{evolvability}. 
Most development teams thus supplement code reviews with {automated} approaches to verification and validation, the most widely-used  of which is \emph{unit testing} \cite{beck2003}. In brief, unit tests are functions that check a single portion of a program against a single well-defined correctness criterion. A suite of such tests can be seen as a partial specification of a program or component.  Test suites complement code review by allowing developers to answer questions like these more easily:
\begin{enumerate}
\item Which functionality has been adequately implemented? What remains to be done?
%\item How much progress does an individual code contribution make? 
\item What modes of usage does the team consider most important? How does the team measure correctness?
\item Does a code contribution cause \emph{regressions} in other parts of a program?
\end{enumerate}
Scientists ask analagous questions:
\begin{enumerate}
\item Which observations are already well-explained by existing models? What is the state-of-the-art? What are the open modeling problems of interest?
%\item How much better is a new model than the state of the art? What are the open problems in my field?
\item What are the contemporary community standards for measuring goodness-of-fit?
\item How do newly-made experimental observations impact the validity of previous models?
\end{enumerate}
But while software engineers can rely on a program's test suite, scientists today must extract this information from a body of scientific publications. This is increasingly difficult. Each publication focuses on just one model and is frozen in time, so it does not consider the latest experimental data or statistical methods. Discovering, precisely characterizing and comparing several models to discover the state of the art and find open modeling problems can often require an encyclopedic knowledge of the literature. Senior scientists can attempt to fill this need by publishing review papers, but in many areas, the number of publications generated every year can be overwhelming \cite{jinha_article_2010}, and comprehensive reviews are published relatively infrequently for any given area. Statisticians commonly complain that scientists are not following best practices and that community standards evolve too slowly because canonical papers or popular reviews advocate outdated methods \cite{grumpystatisticians}. One might compare this process to a software development team relying primarily on carefully-prepared summaries of code review sessions, peppered with cross-references to previous summaries, to understand and validate a complex software project.

This suggests that the scientific process could be improved by the adoption of test-driven methodologies alongside peer review. However, the scientific community presents several unique challenges that must first be addressed:
\begin{enumerate}
\item Unit tests are typically pass/fail, while goodness-of-fit is typically a continuous value (e.g. a $p$-value). 
\item Different communities, groups and individual scientists prefer different goodness-of-fit measures and focus on different sets of observable quantities. This can evolve as, for example, statisticians develop better measures.
\item Model developers compete with one another to predict data, rather than collaborating with one another on a common codebase. They are sometimes reluctant to share certain details of their implementations \cite{badscientists}.
\item Experimental data can itself be erroneous, noisy or inconsistent with other available data.
\item Data formats can differ, sometimes drastically, between experimentalists. Similarly, different modelers may prefer different programming languages or libraries.
\item Professional software developers are typically trained in testing practices and tools, while scientists  rarely have training or experience with testing practices \cite{badscientists2}.
\end{enumerate}
%Anecdotally, one of the greatest fears of new scientists is that a problem that they are working on has already been solved and published.
